### 2018-11-12 source/Ch6-loresEGS/lowres-EGS.tex
asdasda

\pg
In this section, we describe the work done to make a wide-field image of the entire LOFAR primary beam, centred on the Extended Groth Strip (EGS). A large-field, image of this field has never yet been made at such low frequencies. Due to technical limitations, the maximum attainable image size (in terms of number of pixels) will limit resolution to about $5''$. As such, we do not use the international stations for this part of the project, as their contribution would be averaged out within a single pixel. They are only used in later parts of the project, when imaging 3C295 and calibrator sources in the field at high resolutions.
A more complete sky model can be created from this image, and %, allowing for one last round of self-calibration on the 3C295 high-resolution model. 
bright out-of-field sources can be subtracted from high-resolution images of the EGS so that their sidelobe emission do not contaminate the final images. %  This ensures that we do not ``hard-set" errors in our high-resolution model caused by other sources in the field. 
To achieve this, we require direction-dependent calibration. Because we do not use the international LOFAR stations, we do not need a high-resolution model of 3C295.

@@

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP, and consists of a number of 8-hour pointings covering as much of the sky visibile to LOFAR as possible. We analyse a subset of this dataset, a single one of these pointings, an observation performed on the 28th of August 2014 and centred on the EGS. It is an 8-hour-long observation; there exist another four 8-hour pointings in the same direction, but these data are not reduced in this project. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of 110-182 MHz. We use all Core and Remote LOFAR stations. When imaging 3C295 and other sources in the field with the international stations, we only use those which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and British stations).

@@

\pg
We begin with direction-independent calibration to find an estimate of $\Gjones_p\quad\forall p$. This first step is done because this part of the Jones chain is the same throughout the sky, and its inverse solution can therefore be applied directly to the visibilities as follows:
\begin{align}
\Vis_{pq}^\mathrm{corr} &= \Ghatjones_p^{-1} \Vis_{pq} \left( \Ghatjones_q^H \right)^{-1}\\
						&= \Ghatjones_p^{-1}\Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Bmatrix_{s} k_{\Omega,pq} \Ejones_{sq}^H \bigg) \Gjones_q^H \left( \Ghatjones_q^H \right)^{-1}
\end{align}
In so doing, we assume that $\Ejones_{sp}=1quad \forall s,p$ when solving for $\Ghatjones$. In general, this is of course invalid. For the calibrator source, however, we have $\Ejones_{sp} =1 \quad\forall p$ by construction. Because we can choose the calibrator source, the presence of a sufficiently bright source (in our case, an unresolved 3C295) will give extremely good estimations of $\Gjones$. We then have $\Ghatjones_p^{-1}\Gjones_p\sim \I$, and
\begin{align}
\Vis_{pq}^\mathrm{corr} &\sim \sum_{\Omega}\left( \sum_{s\in\Omega} \Ejones_{\Omega p} \Bmatrix_{s} k_{s,pq} \Ejones_{\Omega q}^H \right)\label{eq.vcorr}
\end{align}
and direction-dependent calibration then consists of solving for $\Ehatjones_{\Omega p} = \Ghatjones_p^{-1}\Gjones_p\Ejones_{\Omega p} \sim \Ejones_{\Omega p}$ for all facets -- this means that, if there are errors in the direction-independent gain solutions, they will bias all direction-dependent gain solutions.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefield.DI}.png}}
\caption{\label{image.egs.widefield.noddsc.zoom} LOFAR wide-field image cutout, calibrated wtih the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefiled.DD}.png}}
\caption{\label{image.egs.widefield.doddsc.zoom} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc.zoom} Cutouts of smaller regions of \cref{image.egs.widefield.ddsc}, intended to show that calibration artefacts around fainter sources decrease as well as those around 3C295.}
\end{figure}

@@

\pg
Finally, the quality-based weighting scheme developed in \cref{chapter.paper} was applied throughout this procedure. Noise maps made with and without the weighting scheme are shown in \cref{image.egs.widefield.weights}. A few things are important to note in this comparison. Firstly, the weighting schemes decrease the noise by a factor of $\sim10$: for the noise-map to be visible without weighting schemes, the contrast of the weighted image needed to be pushed way low. For reference, \cref{image.egs.widefield.weighted} is the same image (albeit without the facets shown explicitly) as \cref{fig.EGS.widefield.regions} - though its levels have been changed to better illustrate the effect of the quality-based weighting scheme. This is why we can't see much at all in that image: it is the contrast between \cref{image.egs.widefield.weighted} and \cref{image.egs.widefield.noweights} which is relevant here. The noise distribution is more clearly visible in that figure. Relatedly, by comparing that image and \cref{image.egs.widefield.noweights}, we note that the unweighted image has a different noise per facet, unlike the weighted image. This is due to the fact that a single visibility will be calibrated better or worse depending on the signal-to-noise of individual facets, constrained by the direction-independent calibration which uses the highest signal-to-noise available. It is no coincidence that the facet containing 3C295 is the only one with a noise even remotely comparable to the weighted image's: it is the facet with the best signal-to-noise. This becomes significant when taking into account the fact that the last 2 hours of this observation are much noisier than the first 6: the quality-based weights will decrease their overall contribution \textit{throughout the image}, whereas not using them will lead to different facets having stronger or lesser contributions from these noisy visibilities - and therefore different noise levels. This is also why fringes can clearly be seen in some facets: a single visibility (or a set of a few visibilities) are particularly badly-calibrated, and the associated fringe dominates in the facet. Linking back to \cref{chapter.paper}, because we have a gain solution per visibility \textit{and per facet}, we have one noise-PSF per facet; this suggests that the quality-based weighting scheme could likely be extended to a direction-dependent formalism.

\begin{figure}[h!]
\centering
\begin{subfigure}{.65\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{./image_full_ampphase_di_m.NS_Smooth_noW.noise01.fitsnoisemap.weights.compare}.png}}
\caption{\label{image.egs.widefield.noweights} OFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms, made without using quality-based weights.}
\end{subfigure}
\hfill
\begin{subfigure}{.65\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{./image_full_ampphase_di_m.NS_Smooth.noise01.fitsnoisemap.weights.compare}.png}}
\caption{\label{image.egs.widefield.weighted} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms, made with quality-based weights.}
\end{subfigure}
\caption{\label{image.egs.widefield.weights} Impact of quality-based weighting scheme on the final image.}
\end{figure}

@@

\pg
Another impressive source, EGS-4 was selected for its complex diffuse structure. Like EGS-3, the structure of this source seems to indicate that it would be a radio jet galaxy within a galaxy cluster. Its LOFAR structure matches the lower-resolution NVSS structure, which is a strong indication that it is not spurious. By the same argument as for EGS-3, we expect this diffuse emission to have associated optical emission somewhere along its central axis. As we can see from \cref{fig.egs4.sdss.nvssoverlay}, EGS-4 can be associated to an NVSS source: NVSS J143137+522728. This radio galaxy lies at redshift 0.292 \citepads{2010ApJ...723.1119L}, and is itself associated to the NVSS source in the middle of EGS-4 and its WISE counterpart. SIMBAD has no further association for this radio galaxy.
Further pointings will be required to acquire more information on this source and its environment.

@@

\pg
EGS-5 was chosen due to its size, and was expected to be a very classic radio galaxy. Unfortunately, it lies too close to the local noise to show up in the overlay map (the lowest level of which lies at $5\sigma$). In fact, its lobes were only matched individually. The southern lobe matches to NVSS J142941+510850 \citepads{1997ApJ...475..479W}. The northern lobe matches to NVSS J142954+511147 \citepads{1998AJ....115.1693C}. It is unfortunate that the centre of EGS-5 lies in a region of poor SDSS coverage. Either the original sources are incorrectly identified as individual radio galaxies, and are actually the lobes of a single galaxy, or our analysis of the source is flawed, and we are in fact dealing with two separate sources. The diffuse structure between them in the LOFAR band would seem to indicate that the former case is closer to the truth.

@@

\pg
The work done in this section had the primary goal of creating a decent model of the sky seen by LOFAR's primary beam. To do this, a wide-field, $6''$-resolution image of the fully primary beam was made: this image makes no use of the LOFAR international stations due to technical limitations. Using PyBDSM \citepads{2015ascl.soft02007M}, this model was extracted from an image made after a few rounds of direction-dependent self-calibration.

@@

\pg
Further work on this image is of course both possible and highly desirable: the work done here was artisanal, hand-picking sources and associating them individually to equivalents in other bands. Modern survey work, however, requires automation. Applying Bayesian fitting methods to find most likely counterparts to all the sources extracted in the field could yield useful information, as we would then have access to population statistics. However, doing this would require running yet more self-calibration to remove remaining artefacts, whereas our primary aim is to produce a high-resolution map of the Extended Groth Strip, which has the richest multi-frequency coverage and therefore the most likely to give effective matches to all the sources in the field. We therefore do not proceed to do this work on the low-resolution, wide-field image of the Extended Groth Strip and its neighbourhood, but are likely to return to the data to do this at a later date.

>>>>

\pg
In this section, we describe the work done to make a wide-field image of the entire LOFAR primary beam, centred on the Extended Groth Strip (EGS). A large-field, image of this field has never yet been made at such low frequencies. Due to technical limitations, the maximum attainable image size (in terms of number of pixels) will limit resolution to about $5''$. As such, we do not use the international stations for this part of the project, as their contribution would be averaged out within a single pixel. They are only used in later parts of the project, when imaging 3C295 and calibrator sources in the field at high resolutions.
A more complete sky model can be created from this $\sim5''$-resolution image, and %, allowing for one last round of self-calibration on the 3C295 high-resolution model. 
bright out-of-field sources can be subtracted from high-resolution images of the EGS so that their sidelobe emission do not contaminate the final images. %  This ensures that we do not ``hard-set" errors in our high-resolution model caused by other sources in the field. 
To achieve this, we require direction-dependent calibration. Because we do not use the international LOFAR stations at this point, we do not yet need a high-resolution model of 3C295.

@@

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP, and consists of a number of 8-hour pointings covering as much of the sky visible to LOFAR as possible. We analyse a subset of this dataset, a single one of these pointings, an observation performed on the 28th of August 2014 and centred on the EGS. It is an 8-hour-long observation; there exist another four 8-hour pointings in the same direction, but these data are not reduced in this project. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of 110-182 MHz. We use all Core and Remote LOFAR stations. When imaging 3C295 and other sources in the field with the international stations during \cref{section.3c295,section.decorr}, we only use those which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and British stations).

@@

\pg
We begin with direction-independent calibration to find an estimate of $\Gjones_p\quad\forall p$. This first step is done because this part of the Jones chain is the same throughout the sky, and its inverse solution can therefore be applied directly to the visibilities as follows:
\begin{align}
\Vis_{pq}^\mathrm{corr} &= \Ghatjones_p^{-1} \Vis_{pq} \left( \Ghatjones_q^H \right)^{-1}\\
						&= \Ghatjones_p^{-1}\Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Bmatrix_{s} k_{\Omega,pq} \Ejones_{sq}^H \bigg) \Gjones_q^H \left( \Ghatjones_q^H \right)^{-1}
\end{align}
In so doing, we assume that $\Ejones_{sp}=1\quad \forall s,p$ when solving for $\Ghatjones$. In general, this is of course invalid. For the calibrator source, however, we have $\Ejones_{sp} =1 \quad\forall p$ by construction. Because we can choose the calibrator source, the presence of a sufficiently bright source (in our case, an unresolved 3C295) will give extremely good estimations of $\Gjones$. We then have $\Ghatjones_p^{-1}\Gjones_p\sim \I$, and
\begin{align}
\Vis_{pq}^\mathrm{corr} &\sim \sum_{\Omega}\left( \sum_{s\in\Omega} \Ejones_{\Omega p} \Bmatrix_{s} k_{s,pq} \Ejones_{\Omega q}^H \right)\label{eq.vcorr}
\end{align}
and direction-dependent calibration then consists of solving for $\Ehatjones_{\Omega p} = \Ghatjones_p^{-1}\Gjones_p\Ejones_{\Omega p} \sim \Ejones_{\Omega p}$ for all facets -- this means that, if there are errors in the direction-independent gain solutions, they will bias all direction-dependent gain solutions.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefield.DI}.png}}
\caption{\label{image.egs.widefield.noddsc.zoom} LOFAR wide-field image cutout, calibrated with the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.8\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefiled.DD}.png}}
\caption{\label{image.egs.widefield.doddsc.zoom} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc.zoom} Cutouts of smaller regions of \cref{image.egs.widefield.ddsc}, intended to show that calibration artefacts around fainter sources decrease as well as those around 3C295.}
\end{figure}

@@

\pg
Finally, the quality-based weighting scheme developed in \cref{chapter.paper} was applied throughout this procedure. Noise maps made with and without the weighting scheme are shown in \cref{image.egs.widefield.weights}. A few things are important to note in this comparison. Firstly, the weighting schemes decrease the noise by a factor of $\sim10$: for the noise-map to be visible without weighting schemes, the contrast of the weighted image needed to be pushed way low. For reference, \cref{image.egs.widefield.weighted} is the same image (albeit without the facets shown explicitly) as \cref{fig.EGS.widefield.regions} - though its levels have been changed to better illustrate the effect of the quality-based weighting scheme. This is why we can't see much at all in that image: it is the contrast between \cref{image.egs.widefield.weighted} and \cref{image.egs.widefield.noweights} which is relevant here. The noise distribution is more clearly visible in that figure. Relatedly, by comparing that image and \cref{image.egs.widefield.noweights}, we note that the unweighted image has a different noise per facet, unlike the weighted image. This is due to the fact that a single visibility will be calibrated better or worse depending on the signal-to-noise of individual facets, constrained by the direction-independent calibration which uses the highest signal-to-noise available. It is no coincidence that the facet containing 3C295 is the only one with a noise even remotely comparable to the weighted image's: it is the facet with the best signal-to-noise. This becomes significant when taking into account the fact that the last 2 hours of this observation are much noisier than the first 6: the quality-based weights will decrease the overall contribution of these two hours of data \textit{throughout the image}, whereas not using them will lead to different facets having stronger or lesser contributions from these noisy visibilities - and therefore different noise levels. This is also why fringes can clearly be seen in some facets: a single visibility (or a set of a few visibilities) are particularly badly-calibrated, and the associated fringe dominates in the facet. Linking back to \cref{chapter.paper}, because we have a gain solution per visibility \textit{and per facet}, we have one noise-PSF per facet; this suggests that the quality-based weighting scheme could likely be extended to a direction-dependent formalism.

\begin{figure}[h!]
\centering
\begin{subfigure}{.65\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{./image_full_ampphase_di_m.NS_Smooth_noW.noise01.fitsnoisemap.weights.compare}.png}}
\caption{\label{image.egs.widefield.noweights} OFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms, made without using quality-based weights.}
\end{subfigure}
\hfill
\begin{subfigure}{.65\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{./image_full_ampphase_di_m.NS_Smooth.noise01.fitsnoisemap.weights.compare}.png}}
\caption{\label{image.egs.widefield.weighted} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms, made with quality-based weights.}
\end{subfigure}
\caption{\label{image.egs.widefield.weights} Impact of quality-based weighting scheme on the final image. Note that \cref{image.egs.widefield.weighted} is the same image, albeit at a different contrast.} %%%here%%%
\end{figure}

@@

\pg
Another impressive source, EGS-4 was selected for its complex diffuse structure. Like EGS-3, the structure of this source seems to indicate that it would be a radio jet galaxy within a galaxy cluster. By the same argument as for EGS-3, we expect this diffuse emission to have associated optical emission somewhere along its central axis. As we can see from \cref{fig.egs4.sdss.nvssoverlay}, EGS-4 can be associated to an NVSS source: NVSS J143137+522728.  Its LOFAR structure matches the lower-resolution NVSS morphology, which is a strong indication that this structure is physical. This radio galaxy lies at redshift 0.292 \citepads{2010ApJ...723.1119L}.%, and is itself associated to the NVSS source in the middle of EGS-4 and its WISE counterpart. SIMBAD has no further association for this radio galaxy.
Further pointings will be required to acquire more information on this source and its environment.

@@

\pg
EGS-5 was chosen due to its size, and was expected to be a very classic radio galaxy. Unfortunately, it lies too close to the local noise to show up in the overlay map (the lowest level of which lies at $2\sigma$). In fact, its lobes were only matched individually. The southern lobe matches to NVSS J142941+510850 \citepads{1997ApJ...475..479W}. The northern lobe matches to NVSS J142954+511147 \citepads{1998AJ....115.1693C}. It is unfortunate that the centre of EGS-5 lies in a region of poor SDSS coverage. Either the original sources are incorrectly identified as individual radio galaxies, and are actually the lobes of a single galaxy, or our analysis of the source is flawed, and we are in fact dealing with two separate sources. The diffuse structure between them in the LOFAR band would seem to indicate that the former case is closer to the truth.

@@

\pg
The work done in this section had the primary goal of creating a decent model of the sky seen by LOFAR's primary beam. To do this, a wide-field, $6''$-resolution image of the fully primary beam was made: this image makes no use of the LOFAR international stations due to technical limitations. Using PyBDSM \citepads{2015ascl.soft02007M}, a sky model of the full field of view was extracted from the aforementioned image made after a few rounds of direction-dependent self-calibration.

@@

\pg
Further work on this image is of course both possible and highly desirable: the work done here was artisanal, hand-picking sources and associating them individually to equivalents in other bands. Modern survey work, however, requires automation. Applying Bayesian fitting methods to find most likely counterparts to all the sources extracted in the field could yield useful information, as we would then have access to population statistics. However, doing this would require running yet more self-calibration to remove remaining artefacts, whereas our primary aim is to produce a high-resolution map of the Extended Groth Strip, which has the richest multi-frequency coverage and is therefore the most likely to give effective matches to all the sources in the field. We therefore do not proceed to do this work on the low-resolution, wide-field image of the Extended Groth Strip and its neighbourhood, but are likely to return to the data to do this at a later date.

### 2018-11-12 source/Ch7-3C295/3c295.tex
asdasda

\pg
Because we solve for the full Jones matrix, we cannot apply phase-only corrections\footnote{Indeed, what is the phase of a $2\times2$ matrix of complex numbers?} over multiple self-calibration rounds, as is standard practice. Over self-calibration rounds, we expect that $\alpha$ will change, but that it will remain smooth in frequency. The integrated flux of 3C295 over different passes of self-calibration is shown in \cref{fig.SC.ampfluct}: as this figure shows, the model amplitude can diverge significantly over the self-calibration process. %, sometimes even moving closer to the inital integrated flux, but often moving farther.

% images:
% IMAGES/3c295.fromVLA.6sb.withbeam.pass[1-5].int.model.fits IMAGES/3c295.fromVLA.60sb.withbeam.pass[1-3].int.model.fits IMAGES/MSlist.fullBW.LOBOS6.MSlist.fullBW.txt.uniform.int.model.fits
% script: /home/offler/plotSCamp.py
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/AmpFluct.png}
\caption{\label{fig.SC.ampfluct} Amplitude fluctuation of the models resulting from self-calibration iterations, expressed as a ratio of the model's integrated flux divided by the Scaife \& Healde integrated flux model at the appropriate frequency. Note that the integrated model flux for 324sb-pass1 (the penultimate data point in this plot) was lost due to a careless command: the value shown here is just the average of the two neighbouring values - i.e. those of 60sb-pass3 and 324sb-pass2.}
\end{figure}

@@

\pg
Let us begin by checking whether our core and remote station gains change over time due to our self-calibration procedure. Here, we look at the gains of SB155, on which we solved for gain solutions using our initial and final models. This subband was included in both the 60-subband self-calibration step and the 6-subband self-calibration step.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.firstpass}.png}
\caption{\label{fig.sb155.gains.cs201.before} CS005, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.lastpass}.png}
\caption{\label{fig.sb155.gains.cs201.after} CS005, last calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.RS305.before} RS305, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.RS305.afer} RS305, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.CR} Gain amplitude \& phase for CS201 of SB155}
\end{figure}

@@

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes were found to compensate for this effect quite well, and minimise the associated image deterioration\footnote{The evidence for this is in the previous chapter, as the test data is the same as used here; nevertheless, an extra test was carried out in the context of these calibration and imaging tests for the sake of completeness, and found that quality-based weighting schemes resulted in better results than simply flagging the last 2 hours.}. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602 of SB155}
\end{figure}

@@

\pg
We see that some of the structure in the amplitude gains shown in \cref{fig.fullBW.DE601HBA} goes away over the course of self-calibration, though one of the two arcs present in the plot appears to decrease in amplitude. For \cref{fig.fullBW.DE602HBA}, we see that the pattern in the amplitude plot clearly changes - the fact that the pattern in \cref{fig.fullBW.DE602HBA.before} does not appear to be either dispersion or scintillation pattern would indicate that it consists of unmodeled flux. While some structure remains in the last 2 hours and low frequencies for this antenna's gain solution amplitudes, the last 2 hours of the observation are generally poor compared to the first six: we expect that the quality-based weighting scheme will minimise the impact of these poor calibration solutions on the final images. In \cref{fig.fullBW.DE603HBA}, we see that the self-calibration process was clearly less effective than it was for DE602HBA: while some of the structure was minimised (in the first two hours particularly: we go from the peculiar ``ellipse" amplitude pattern to something which looks more like scintillation), a lot clearly remains. Our self-calibration process could therefore likely be taken further. However, in our last step, we use 324 out of the full 366 subbands (leaving some out due to poor quality); a single round of self-calibration with this full dataset takes 5000 minutes on the best computing cluster available for this work. Because a single round of self-calibration would, at this point, take as much time as imaging a portion of the EGS due to computational constraints, we treat the self-calibration process as ``having converged" at this point: while it could be yet improved, we have reached a noise of 2.78 mJy.

\clearpage
% % % % % % % % % % % % % % % % % % % % % % %
\section{Overlays \& Spectral Analaysis}

@@

\pg
All the same, this project served as a useful exercise in both developing a good working methodology and in testing calibration strategies for international LOFAR datasets. Much of the development which went into this project and led to no directly useful results still proved an extremely helpful context in which to understand how to improve the LOFAR pipeline. In particular, gain smoothing\footnote{i.e. the application of a smoothing operation in gain phase and amplitude, justified by the fact that they are expected to be smooth and so jagged behaviour is likely erroneous.} has shown promising result on LOTSS images, but was not used in this project because of the full-Jones solutions: properly implementing this smoothing in a framework capable of accounting for polarisation could provide the kind of regularisation that our gain solutions need.

>>>>

\pg
Because we solve for the full Jones matrix, we do not apply phase-only corrections over multiple self-calibration rounds, as is standard practice. This is because both the phase and amplitude of the gains need to be solved for as a function of time. Over self-calibration rounds, we expect that $\alpha$ will change, but that it will remain smooth in frequency. The integrated flux of 3C295 over different passes of self-calibration is shown in \cref{fig.SC.ampfluct}: as this figure shows, the model amplitude can diverge significantly over the self-calibration process. %, sometimes even moving closer to the inital integrated flux, but often moving farther.

% images:
% IMAGES/3c295.fromVLA.6sb.withbeam.pass[1-5].int.model.fits IMAGES/3c295.fromVLA.60sb.withbeam.pass[1-3].int.model.fits IMAGES/MSlist.fullBW.LOBOS6.MSlist.fullBW.txt.uniform.int.model.fits
% script: /home/offler/plotSCamp.py
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/AmpFluct.png}
\caption{\label{fig.SC.ampfluct} Amplitude fluctuation of the models resulting from self-calibration iterations, expressed as a ratio of the model's integrated flux divided by the Scaife \& Healde integrated flux model at the appropriate frequency. Note that the integrated model flux for 324sb-pass1 (the penultimate data point in this plot) was unfortunately lost due to a careless command: the value shown here is just the average of the two neighbouring values - i.e. those of 60sb-pass3 and 324sb-pass2.}
\end{figure}

@@

\pg
Let us begin by checking whether our core and remote station gains change over time due to our self-calibration procedure. Here, we look at the gains of SB155, on which we solved for gain solutions using our initial and final models. This subband was included in both the 60-subband self-calibration step and the 6-subband self-calibration step.
\begin{figure}[h!]
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.firstpass}.png}
\caption{\label{fig.sb155.gains.cs201.before} CS005, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.CS005HBA0.lastpass}.png}
\caption{\label{fig.sb155.gains.cs201.after} CS005, last calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.RS305.before} RS305, first calibration}
\end{subfigure}
\begin{subfigure}{.9\textwidth}
\includegraphics[width=\linewidth]{images/{SB155.RS305HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.RS305.afer} RS305, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.CR} Gain amplitude \& phase for stations CS201 and RS305 of SB155. These are a core and remote stations, respectively.}
\end{figure}

@@

\pg
As for the international station gains, they are shown below for three of the seven international stations used for our observation\footnote{Namely DE602, SE607 and UK608 - these are a German, Swedish and British station respectively.}, in \cref{fig.sb155.gains.de602,fig.sb155.gains.se607,fig.sb155.gains.uk608}. We show only one of the four German stations. As we can see, while structure very clearly remains in some of these amplitude curves (notable in the last two hours, e.g. \cref{fig.sb155.gains.DE602.after}), some of the structure disappears from these amplitude curves. It should be noted that the last two hours of the observation are generally much noisier than the first 6; the last quarter of these plots show this. Quality-based weighting schemes were found to compensate for this effect quite well, and minimise the associated image deterioration\footnote{The evidence for this is in the previous chapter, as the test data is the same as used here; nevertheless, an extra test was carried out in the context of these calibration and imaging tests for the sake of completeness, and found that quality-based weighting schemes resulted in better results than simply flagging the last 2 hours.}. Note that the phase wraps extremely fast in the international station gain plots: this is because phase is calculated relative to CS001HBA0, the first antenna in our array. Because the international stations are much farther away from this core station than other core or remote stations are, we expect the associated phase solutions to wrap faster.
\begin{figure}[h!]
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.firstpass}.png}
\caption{\label{fig.sb155.gains.DE602.before} DE602, first calibration}
\end{subfigure}
\begin{subfigure}{.9\linewidth}
\includegraphics[width=\linewidth]{images/{SB155.DE602HBA.lastpass}.png}
\caption{\label{fig.sb155.gains.DE602.after} DE602, last calibration}
\end{subfigure}
\caption{\label{fig.sb155.gains.de602} Gain amplitude \& phase for DE602 of SB155}
\end{figure}

@@

\pg
We see that some of the structure in the amplitude gains shown in \cref{fig.fullBW.DE601HBA} goes away over the course of self-calibration, though one of the two arcs present in the plot appears to decrease in amplitude. For \cref{fig.fullBW.DE602HBA}, we see that the pattern in the amplitude plot clearly changes - the fact that the pattern in \cref{fig.fullBW.DE602HBA.before} does not appear to be either dispersion or scintillation pattern would indicate that it consists of unmodeled flux. While some structure remains in the last 2 hours and low frequencies for this antenna's gain solution amplitudes, the last 2 hours of the observation are generally poor compared to the first six: we expect that the quality-based weighting scheme will minimise the impact of these poor calibration solutions on the final images. In \cref{fig.fullBW.DE603HBA}, we see that the self-calibration process was clearly less effective than it was for DE602HBA: while some of the structure was minimised (in the first two hours particularly: we go from the peculiar ``parabolic" amplitude pattern of \cref{fig.fullBW.DE603HBA.before} to something which looks more like scintillation), a lot clearly remains. Our self-calibration process could therefore likely be taken further. However, in our last step, we use 324 out of the full 366 subbands (leaving some out due to poor quality); a single round of self-calibration with this full dataset takes 5000 minutes on the best computing cluster available for this work. Because a single round of self-calibration would, at this point, take as much time as imaging a portion of the EGS due to computational constraints, we treat the self-calibration process as ``having converged" at this point: while it could be yet improved, we have reached a noise of 2.78 mJy/bm.

\clearpage
% % % % % % % % % % % % % % % % % % % % % % %
\section{Overlays \& Spectral Analysis}

@@

\pg
All the same, this project served as a useful exercise in both developing a good working methodology and in testing calibration strategies for international LOFAR datasets. A lot of the work which went into this project led to dead ends, in the sense that they gave no directly useful results - but they still proved to be an extremely helpful context in which to understand how to improve the LOFAR pipeline. In particular, gain smoothing\footnote{i.e. the application of a smoothing operation in gain phase and amplitude, justified by the fact that they are expected to be smooth and so jagged behaviour is likely erroneous.} has shown promising result on LOTSS images, but was not used in this project because of the full-Jones solutions: properly implementing this smoothing in a framework capable of accounting for polarisation could provide the kind of regularisation that our gain solutions need.
